% CBSLAO -- Cost- and SLA-Bounded Orchestration for LLM-Agent Tool/Service Composition
% Target venue: ICSOC 2026 (research track, Springer LNCS)
% Draft v0.2 -- sections are modular under ./sections/.
%
% Build: pdflatex main ; bibtex main ; pdflatex main ; pdflatex main
% (Assumes llncs.cls and splncs04.bst are on TEXINPUTS.  See README.md.)
\documentclass[runningheads]{llncs}

% -------------------- Encoding & fonts --------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}

% -------------------- Math & symbols --------------------
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{amsthm}
\usepackage{mathtools}

% -------------------- Floats, graphics, tables --------------------
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}

% -------------------- Algorithms --------------------
\usepackage{algorithm}
\usepackage{algpseudocode}

% -------------------- Links, colour, misc --------------------
\usepackage{xcolor}
\usepackage[colorlinks=true,
            linkcolor=blue!60!black,
            citecolor=blue!60!black,
            urlcolor=blue!60!black]{hyperref}
\usepackage{url}

% -------------------- Theorem-like environments --------------------
% llncs defines \theorem, \lemma, \proposition, \corollary, \remark, \definition.
% We add a conjecture environment.
% \spnewtheorem{conjectureenv}[theorem]{Conjecture}{\bfseries}{\itshape}
% \newenvironment{conjecture}{\begin{conjectureenv}}{\end{conjectureenv}}

% -------------------- Author macros --------------------
\newcommand{\CBSLAO}{\textsc{CBSLAO}}
\newcommand{\CBUC}{\textsc{CBUC}}
\newcommand{\BwK}{\textsc{BwK}}
\newcommand{\ReAct}{\textsc{ReAct}}
\newcommand{\Oh}{\mathcal{O}}
\newcommand{\tOh}{\widetilde{\mathcal{O}}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\DeclareMathOperator*{\argmax}{arg\,max}

% -------------------- Title, authors --------------------
\title{Cost- and SLA-Bounded Orchestration for LLM-Agent Tool/Service Composition}
\titlerunning{CBSLAO}

\author{Anonymous\inst{1}}
\authorrunning{Anon.~et~al.}
\institute{Anonymous Institution\\
\email{anon@example.org}}

% ============================================================
\begin{document}
\maketitle

\begin{abstract}
LLM-based agents increasingly orchestrate external tools and services
to complete user tasks. A central but under-analysed problem is that
mainstream agent controllers---\ReAct~loops, LangChain's
\texttt{AgentExecutor}, the OpenAI Agents SDK, and AutoGen---invoke
tools without formal cost or latency guarantees, routinely exceeding
user-declared token budgets and service-level objective (SLO) deadlines.
We formalise the \emph{Cost- and SLA-Bounded Agent Orchestration}
(\CBSLAO) problem, in which an agent selects an adaptive sequence of
tool invocations maximising expected utility subject to
chance-constrained budget and deadline requirements. We prove that
\CBSLAO~is NP-hard by reduction from 0/1 Knapsack and, more strongly,
$(1 - 1/e + \varepsilon)$-inapproximable by reduction from
Max-$k$-Coverage---a bound matched (up to $\varepsilon$) by a
$(1-1/e)$-approximation in the offline non-adaptive
monotone-submodular case. An explicit construction shows the
adaptivity gap is at least $4/3$ on a worst-case family and at most
an absolute constant on every instance. We then present \CBUC~(\emph{Cost-Budgeted Upper Confidence}), an
online algorithm that combines schema-type-based offline pruning with a
budget-aware UCB selector. Under bounded sub-Gaussian cost and utility
noise, \CBUC~attains $\tOh(\sqrt{KT})$ regret against the best
feasible non-adaptive policy for $K$ arms over $T$ rounds, via a
coupled multivariate empirical-Bernstein concentration that saves a
constant factor over a decoupled analysis.
In an extended controlled simulator spanning $18$ workload cells,
$20$ seeds per cell, and $11$ policies ($3{,}960$ replicate rows),
\CBUC~cuts budget-overrun from $36.9\%$ (\ReAct-capped) and $28.9\%$
(\BwK-vanilla) to $1.1\%$ and eliminates all deadline misses. This
safety gain is not free: \CBUC's utility is $0.587$, compared with
$0.806$ for a Lagrangian primal-dual baseline that still overruns
budget in $23.1\%$ of runs. We therefore position \CBUC~as a
conservative governance policy for settings where budget/SLO
violations are expensive, not as a universally utility-optimal
controller. All headline results are reproducible from a released
simulator; no proprietary LLM API is required.
\keywords{LLM agents \and service composition \and
bounded rationality \and online learning \and chance constraints \and
service-oriented computing.}
\end{abstract}

\input{sections/introduction}
\input{sections/related_work}
\input{sections/formulation}
\input{sections/hardness}
\input{sections/algorithm}
\input{sections/evaluation}
\input{sections/threats}
\input{sections/limitations}
\input{sections/repro}
\input{sections/conclusion}

% -------------------- Bibliography --------------------
\bibliographystyle{splncs04}
\bibliography{references}

% -------------------- Appendix --------------------
\appendix
\input{sections/appendix}

\end{document}
