\section{Limitations and Future Work}
\label{sec:limitations}

\begin{itemize}
  \item \CBSLAO's regret result assumes Markovian cost and utility;
    real agent controllers depend on full conversational history, tool
    descriptions, prior outputs, and scratchpad context. A
    contextual-bandit or meta-bandit extension is required before
    positioning \CBUC~as a drop-in production controller.
  \item \CBUC~treats each tool as a single arm; tools with internal
    state (e.g., stateful retrievers) may require a contextual-bandit
    extension.
  \item Type-closure pruning assumes hard schema typing. Production
    registries with fuzzy or partial schemas need probabilistic type
    matching and calibrated abstention, otherwise pruning may either
    admit nearly every tool or remove valid ones.
  \item We do not address \emph{adversarial} tool providers; composition
    with adversarial discovery (cf.\ prompt-injection in tool
    descriptions) is an open problem.
  \item Our hardness result is against optimal utility; fine-grained
    hardness of approximation is open.
  \item Only the CC-knapsack baseline explicitly models the
    \emph{non-adaptive} optimum that \CBUC's regret bound targets.
    Tight instance-adaptive regret against an adaptive oracle remains
    open; \cite{dean2008stochastic}-style adaptivity-gap bounds are
    an avenue.
\end{itemize}

\paragraph{Open problems.}
Two theoretical questions raised by the present analysis are left
open. (O1) \emph{Instance-adaptive regret against the adaptive oracle.} Our
bound targets $\pi^\dagger_{\mathrm{CC}} \in \Pi_{\mathrm{non}}$.
Empirically (Remark~\ref{rem:adaptivegap}) the non-adaptive
oracle is within $\approx 1\%$ of the informed-knapsack upper bound on
our instances, suggesting the adaptivity gap for two-resource CC
stochastic knapsack is small, but a distribution-free, tight regret
bound against the adaptive oracle is not known. (O2)
\emph{Beyond sub-Gaussianity.} Corollary~\ref{cor:cvar} shows that
CVaR-based feasibility preserves the regret rate up to $\sqrt{\log T}$,
but the constant-factor comparison between the two corrections under
Pareto-$\alpha \in (1, 2]$ cost is not pinned down.
