\section{Related Work}
\label{sec:related}

We survey four bodies of work most directly related to \CBSLAO.

\paragraph{LLM agents and tool use.}
\ReAct~\cite{yao2023react}, Toolformer~\cite{schick2023toolformer},
Chameleon~\cite{lu2023chameleon}, and production agent frameworks
(OpenAI Agents SDK, LangChain \texttt{AgentExecutor}, AutoGen) establish
the controller archetype. These systems optimise task success; none
offers formal cost or latency guarantees.

\paragraph{Tool-use benchmarks.}
ToolBench~\cite{qin2023toolllm}, $\tau$-bench~\cite{yao2024taubench},
and the Berkeley Function-Calling
Leaderboard~(BFCL)~\cite{bfcl2024} provide task traces but do not
expose cost/latency ground truth in a form directly reusable for
chance-constrained optimisation.

\paragraph{Service-oriented composition under QoS.}
Classical QoS-aware service composition~\cite{zeng2004qos,alrifai2009qos}
formulated composition as constrained optimisation; \CBSLAO~inherits
this tradition but must additionally contend with the
\emph{non-stationary} and \emph{stochastic} cost of LLM-mediated
calls.

\paragraph{Stochastic knapsack with deadlines.}
Dean, Goemans \& Vondr\'{a}k~\cite{dean2008stochastic} and followups
give adaptive/non-adaptive gap results for stochastic knapsack. We
adopt their framing for the adaptivity-gap remark in
\S\ref{sec:formulation}.

\paragraph{Budget-aware bandits.}
Badanidiyuru, Kleinberg \& Slivkins~\cite{bwk2018} on
\emph{bandits with knapsacks} (\BwK) gives the closest existing online
framework; \CBUC~specialises \BwK~to the LLM-agent setting with
schema-type pruning and explicit chance constraints.

\paragraph{Constraint-aware decision baselines.}
Several strong families of constrained-decision algorithms are
deployable competitors: \emph{primal-dual Lagrangian} methods for
constrained bandits and constrained MDPs~\cite{mahdavi2012tradeoff,agrawal2014bwk}, \emph{tail-aware} formulations that replace the
sub-Gaussian concentration with CVaR~\cite{tamkin2020cvar}, and
offline \emph{chance-constrained knapsack}
solvers~\cite{kleinberg2000stochasticpacking} that inflate costs by a
safety margin and solve 0/1 knapsack once. We include a representative
of each in our evaluation (\S\ref{sec:evaluation}).

